Preventing a Nuclear ISIL: \\Recommendations for United States Policy

### 2015-05-11 draft.tex
Added in subsection 6.6.3 - "Global Threat Reduction Initiative."

Existing Policy: \textcolor{red}{Target ISIL Income Stream Proposed Policy}

@@

STILL MISSING

@@

Contributing Agency: DNDO, PNNL, Bechtel Nevada \normalfont

@@

Existing Policy:  \textcolor{red}{IAEA Monitoring}

Proposed Policy:  \textcolor{red}{Push for Additional Protocol signatures}

Lead Agency: DoS

@@

Proposed Policy:  \textcolor{red}{Renew and broaden GTRI commitment}

@@

GTRI existing policy :

GTRI prevents nuclear material from getting into the wrong hands by: removing material from improperly protected places, converting nuclear  reactors from highly enriched uranium (HEU) to low enriched uranium (LEU), and improving security at nuclear facilities that are vulnerable to theft or sabotage [1]. Thus far, 49 HEU research reactors have been converted or shut down. Currently, security enhancement involves installation of detectors, such as Passive Infrared Sensors (PIR), Remote Monitoring System (RMS), Personal Radiation Detector (PRD), etc [2]. Courses have been developed to train personnel of nuclear facilities on how to use the advanced detectors and equipment to secure their nuclear facilities, as well as how to respond in case of potential theft or sabotage. GTRI should provide travel funding for people to attend courses where they can get a theoretical basis and hands-on security experience. GTRI is also collaborating with universities to create a nuclear security curriculum.

Proposed policy

@@

The GTRI should put emphasis on establishing collaboration with international organizations and countries throughout the world to enhance nuclear security.  In addition to working directly in foreign countries, GTRI should provide training and equipment to local state agencies to enable them to assure the safety of their nuclear facilities.

Russia still houses the most radioactive material and has dozens of nuclear reactors that still use HEU fuel.  Bilateral agreement with countries like Russia should be established in addition to the GTRI to allow close adjustment to the specific situation.

Budget and Implementation

Converting or possibly shutting down the 200 HEU research reactors is expected to be completed by 2035, under the current budget plan.  By 2018, 98\% of excess and vulnerable weapon grade material will be removed[3]. 8,500 buildings with will be secured with high-priority nuclear and radiological material is predicted to be finished in 2044[4]. The budget for fiscal years 2012, 2013, and 2014 are \$503453, \$501048 and \$424487.

% Sources
% [1] GTRI: Reducing Nuclear Threats, National nuclear security administration, http://nnsa.energy.gov/mediaroom/factsheets/reducingthreats
% [2] Global Threat Reduction Initiative, nuclear and radiological material protection,  NNSA fact sheet, http://osrp.lanl.gov/Documents/NNSAGTRIfact%20sheet.pdf
% [3] Department of Energy FY 2014 Congressional Budget Request, DOE/CF-0084, Volume 1, National Nuclear Security Administration Office of the Administrator Weapons Activities Weapons Activities Defense Nuclear Nonproliferation Naval Reactors, April 2013
% [4] US Budget Request Shows Eroding Focus on Nuclear Security, Nickolas Roth, March 25, 201l

@@

Proposed Policy:  \textcolor{red}{????????}

@@

Contributing Agency: ??????????  \normalfont

@@

Existing Policy: \textcolor{red}{Attribution through nuclear forensics}

Proposed Policy: \textcolor{red}{Declaratory policy regarding use of nuclear forensic attribution}

Lead Agency: DoD, DOS

Contributing Agency: DNDO, IC, NNSA, DTRA \normalfont

@@

Proposed Policy: \textcolor{red}{Continue to target necessary infrastructure}

>>>>

Existing Policy: \textcolor{red}{Target ISIL Income Stream}

@@

STILL MISSING!!!

Eric Harvey, your assigned section is still missing after being 3 weeks late! Get it to us yesterday!

@@

Contributing Agency: DNDO, DTRA, DOS, Bechtel Nevada  \normalfont

@@

Existing Policy:  \textcolor{red}{IAEA Monitoring \& the Additional Protocol (AP)}

Proposed Policy:  \textcolor{red}{Push for AP signatures from high risk countries (Egypt, Syria, \& Pakistan)}

Lead Agency: DoS, ISN

@@

Proposed Policy:  \textcolor{red}{Renew and broaden GTRI commitment; emphasize international  collaboration }

@@

% GTRI existing policy :

The Global Threat Reduction Initiative (GTRI) prevents nuclear material from getting into the wrong hands by removing SNM from improperly protected facilities, converting nuclear  reactors from highly enriched uranium (HEU) to low enriched uranium (LEU), and improving security at nuclear facilities that are vulnerable to theft or sabotage \cite{NationalNuclearSecurityAdministration2014}. Thus far, 49 HEU research reactors have been converted or shut down. Currently, security enhancement involves installation of detectors, such as Passive Infrared Sensors (PIR), Remote Monitoring Systems (RMS), and Personal Radiation Detectors (PRD) \cite{NationalNuclearSecurityAdministrationa}. Courses have been developed to train personnel of nuclear facilities on how to use the advanced detectors and equipment to secure their nuclear facilities, as well as how to respond in case of potential theft or sabotage. GTRI should provide travel funding for people to attend courses where they can get a theoretical basis and hands-on security experience. GTRI is also collaborating with universities to create a nuclear security curriculum.

% Proposed policy

@@

The GTRI should put emphasis on establishing collaboration with international organizations and countries throughout the world to enhance nuclear security.  In addition to working directly in foreign countries, GTRI should provide training and equipment to local state agencies, to enable them to assure the safety of their nuclear facilities.

Russia still houses the largest quantities of radioactive material and has dozens of nuclear reactors that still use HEU fuel.  Bilateral agreement with countries such as Russia should be established in addition to the GTRI to allow close adjustment to the specific situation.

% Budget and Implementation

Converting or possibly shutting down the 200 HEU research reactors is expected to be completed by 2035, under the current budget plan.  By 2018, 98\% of excess and vulnerable weapons-grade material will be removed \cite{Position2014}. 8,500 buildings   with high-priority nuclear and radiological material will be secured, predicted to be finished in 2044 \cite{Roth2014}. The budget for fiscal years 2012, 2013, and 2014 are \$503,453, \$501,048 and \$424,487, respectively.

@@

Proposed Policy:  \textcolor{red}{Strengthen existing bilateral arrangements; Engage Iran and North Korea}

@@

Contributing Agency: DoD, DoE  \normalfont

@@

Existing Policy: \textcolor{red}{Nuclear Forensics Attribution Act}

Proposed Policy: \textcolor{red}{Declaratory policy on use of force for nuclear technology transfer}

Lead Agency: DOS, DNDO/NTNFC

Contributing Agency: DoD, IC, NNSA, DTRA \normalfont

@@

Proposed Policy: \textcolor{red}{Target natural resources; protect existing infrastructure}
